---
title: "Assignment II: Group B"
author: "LI, Yuan"
format: 
  html:
    toc: true
    code-fold: false
    embed-resources: true # <= check here: https://quarto.org/docs/reference/formats/html.html
execute:
  echo: true    # <= knitr::opts_chunk$set(echo = TRUE) in Rmarkdown
  warning: false
  message: false
---

# Name

[**Name**]{style="color:red"}**: LI Yuan**

# Generative AI declaration

Students may use generative AI tools during assignment, for example, to check and improve code, provided that the content is original and reflects the student's own contributions. If generative AI is used, students must declare its use in the box below. An example statement is: “*During the assignment, I used \[XXXX\] to \[XXXX\]. I subsequently reviewed and edited the code and take(s) full responsibility for the final version of the code.”*

::: callout
During the assignment, I used \[Deepseek\] to \[adjust parameters\]. I subsequently reviewed and edited the code and take(s) full responsibility for the final version of the code.
:::

# Dataset

Each group member may choose either of the following dataset:

**Built-in dataset**

-   `palmerpenguins::penguins`
-   `ggplot2::msleep`
-   `ggplot2::midwest`
-   `datasets::quakes`

**Public dataset**

BIOTIME (CC0) <https://biotime.st-andrews.ac.uk/usageGuidelines.php>

-   `BIOTIME_raw_data_192.csv` (uploaded on Canvas) (Whale data from <https://biotime.st-andrews.ac.uk/selectStudy.php?study=192>)
-   `BIOTIME_raw_data_232.csv` (uploaded on Canvas) (Fish data from <https://biotime.st-andrews.ac.uk/selectStudy.php?study=232>)

# Figures by each group member

### Dataset

::: callout-important
[Which dataset did you choose?]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

ggplot2::midwest
:::

### Introduction

::: callout-important
[What is the dataset about? What variables are included in the data? How was the data taken/measured?]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

This dataset contains demographic, education level and poverty level information of 437 observations and 28 variables in the Midwest United States (Illinois, Indiana, Michigan, Ohio, Wisconsin) from the 2000 US Census. Below are the  variables information:

- `PID`: Unique county identifier (categorical)
- `county`: Name of the county (categorical)
- `state`: Name of the state (categorical)
- `area`: Area of county (units unknown) (float)
- `poptotal`: Total population (number)
- `popdensity`: Population density (person/unit area) (float)
- `popwhite/popblack/popamerindian/popasian/popother`: The Percent of people of different races
- `percwhite/percblack/percamerindan/percasian/percother`: The number of people of different races with (%) (float)
- `popadults`: Number of adults (number)
- `perchsd/percollege/percprof/percasian`: The percent of different education level with(%) (float)
- `poppovertyknown`: Population with known poverty status (numeric)
- `percpovertyknown`: Percent of population with known poverty status (float)
- `percbelowpoverty`: Percentage of population below poverty line (float)
- `percchildbelowpovert/percadultpoverty/percelderlypoverty`: Percent of children/adults/adults below poverty line (float)
- `inmetro`: County considered in a metro area (boolean)
- `category`: County considered in a metro area (categorical)


The data is derived from the US Census Bureau's decennial census (primarily the 2000 Census), with county-state level estimates for demographic, educational, and poverty characteristics. 
:::

### Interactive plot(s) using `plotly`

::: callout-important
[Please explain which plot type you choose and why. Also, please discuss the strengths and weaknesses of your plot(s).]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

I chose an interactive grouped scatter plot (bubble plot) with linear trend lines, implemented via ggplot2 + plotly. Below are the reasons.

strengths:

(1) A scatter plot can show the bivariate relationship between two continuous variables.
(2) muitivariate narrative, can add a third  dimension. Mapping a specific variable to point size for adds a third continuous variable.
(3) interactivity, hover info/highlighting.
(4) Can add grouped linear regression lines (geom_smooth) and color.

weaknesses:

(1) The data points are too densely packed. Can obscure individual data points, even with transparency.
(2) Size interpretation ambiguity.
(3) Trend line limitations. Linear regression assumes a linear relationship, which may not perfectly capture the true, potentially non-linear, association between two variable.
:::

Note: At least one plot is required, but multiple plots are allowed

::: callout-important
[What is the purpose of this visualization?]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

Explore the relationship between educational attainment (high school graduation rate) and poverty rate in counties of the U.S. Midwest, while stratifying the analysis by three distinct education metrics (high school diploma, college degree, and professional degree) to compare their differential associations with poverty. It also incorporates county population size as an additional dimension to examine scale effects.

Test the core hypothesis: Examine whether higher educational attainment is associated with lower poverty rates at the county level, and quantify how this relationship varies across different education levels.

- Bubble Plot:  Directly compare the education-poverty correlation for three education level to identify gradient effects.
- Large vs. Small Counties: Bubble size = total population, to show differences in county scale.
- Color = Education level: Compares the education-poverty relationship.
- Trend Lines = Grouped Linear Regression: Compare the strength of the negative correlation across the three.
- Interactive Hover: Displays county name + detailed data, for exploring specific details.
- Highlight Filter: Focuses on one category at a time, to quickly compare specific education groups patterns.
:::

Note: If you want to specify setting for a code chunk, you can use `#|` for each option (for example, `#| echo: true`). Rmarkdown code chunk still works (for example, `{r echo = true}`), but Quarto style is recommended.

```{r}
#| echo: true
# YOUR CODE HERE

library(ggplot2)
library(plotly)
library(dplyr)
library(tidyr)

# Bubble chart showing interaction among three education level groups
# Changes in poverty rate as education level decreases, accompanied by the corresponding population size

# 1.  Data Preprocessing: convert wide format (perchsd/percollege/percprof) to long format
# generate edu_type (education category) and edu_value (corresponding percentage) columns for grouping
data = midwest %>%
  pivot_longer(cols = c(percprof, percollege, perchsd), # Three education variables for grouping
               names_to = "edu_type",     # Education category (grouping basis)
               values_to = "edu_value") %>%     # Corresponding percentage values
  mutate(edu_type = factor(edu_type,     # Add readable labels
      levels = c("percprof", "percollege", "perchsd"),
      labels = c("Professional", "College", "High School")))

# 2. Bind shared key
midwest_key = highlight_key(data, ~edu_type)

# 3. Initialize Shaped ggplot scatter plot
Shaped_plot = 
  ggplot(midwest_key, aes(x = edu_value, y = percbelowpoverty, 
                      color = edu_type, 
                      shape = edu_type)) +
  
  # 4. Add Shaped scatter point layer 
  geom_point(alpha = 0.6, 
             aes(size = poptotal, 
                 text = paste("<b>County: ", paste(county, "-", state), "</b>\n",
                              "-----------------------------------------\n",
                              "Education Type: ", edu_type, "\n",
                               "Education Rate: ", round(edu_value, 1), "%\n",
                              "Poverty Rate: ", round(percbelowpoverty, 1), "%\n",
                              "Population: ", format(poptotal, big.mark = ",")))) +
  
  # 5. Add grouped linear trend lines + confidence interval
  geom_smooth(aes(group = edu_type),    # Plot three trend lines grouped by Metro status
              method = "lm", se = TRUE,       # Linear regression, 95% CI 
              linewidth = 1, alpha = 0.3,
              show.legend = FALSE) +
  
  # 6. optimize chart
  scale_size_continuous(name = "Total Population", range = c(2, 15)) +    # Point size range (min/max in pixels)）
  scale_color_manual(values = c("skyblue", "orange", "pink"), name = "Education Type") +
  scale_shape_manual(values = c(24,25,23), name = "Education Type") +
  scale_x_continuous(breaks = seq(0, 100, by = 10),  # major ticks: 0,10,20,...,100
                     limits = c(0, 100)) +    # x-axis range: 0-100%
  labs(title = "Education Level vs. Poverty in Midwest US",
       x = "Education Attainment Rate(%)", y = "Poverty Rate (%)") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14, face = "bold",  hjust = 0.5),
        axis.title = element_text(size = 10),
        axis.text = element_text(size = 8),
        panel.border = element_rect(),     # add outer border
        axis.ticks = element_line(),     # add x and y axis ticks
        legend.position = "right",
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8))

# 7. convert to plotly + enable category hover highlight
ggplotly(Shaped_plot, tooltip = "text", height = 400) %>%
  highlight(on = "plotly_hover",      # Trigger highlight on mouse hover
            off = "plotly_doubleclick",  
            dynamic = TRUE,      # Dynamic highlight update
            persistent = FALSE,          # Keep highlight after mouse move away
            selectize = TRUE,     
            color = c("white", "purple", "red2", "blue3")) %>%
  layout(margin = list(t = 80),     # Expand top space to avoid title occlusion
         xaxis = list(showgrid = TRUE, minor = list(ticks = "inside", showgrid = FALSE, dtick = 5)), # Show minor ticks
         yaxis = list(showgrid = TRUE, minor = list(ticks = "inside", showgrid = FALSE, dtick = 5)))    
```


### Animation using `gganimate`

::: callout-important
[Please explain which animation type you choose and why. Also, please discuss the strengths and weaknesses of your animation.]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

I chose a state-transition animated grouped scatter plot  implemented via ggplot2 + gganimate, using transition_states() to sequentially display data by education level. Below are the reasons: 

(1) Supports categorical grouping.
(2) Shadow_mark(): Can  preserves full data context, convenient for comparing data.
(3) Can creates a smooth, intuitive flow. 

Strengths: 

(1) An orderly narrative perspective.
(2) Support cumulative static background images and record the data of the previous step.
(3) Multivariate display: point color/shape, animation state grouping, and continuous variables to axes, effectively visualizing 4 dimensions of data.
(4) Trend line integrate: show how the trend line changes.

Weaknesses:

(1) The transition is too fast, making it easy to overlook some important details.
(2) There is too much data, and the legends overlap, even with transparency.
(3) The GIF file is too large, and it takes a long time to load each time.
:::

Note: At least one animation is required, but multiple animations are allowed

::: callout-important
[What is the purpose of this visualization?]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

Visualization how the relationship between educational attainment and poverty rate changes across three distinct education levels (High School, College, Professional Degree) for counties in 5 U.S. Midwest states.


Test the core hypothesis: Examine whether higher educational attainment is consistently associated with lower poverty rates at the county level, and quantify how the strength of this negative correlation varies across high school, college, and professional degree attainment.
State-level comparison: Use color/shape encoding to distinguish 5 Midwest states. Comparing how the education-poverty relationship differs geographically.


- Color/Shape: Use to distinguish 5 Midwest states. Comparing how the education-poverty relationship differs geographically.
- Animation State: Education level, sequentially displaying high school, college, and professional degree groups.
- Linear Trend Lines: Grouped regression lines to quantify the strength of the negative correlation for each education level.
:::

```{r}
# YOUR CODE HERE

library(gganimate)

# Animation showing three education level groups
# Changes in poverty rates in each state as education levels decrease

# 1. Set up the plot and mapping
animated_plot = ggplot(data, aes(x = edu_value, y = percbelowpoverty, color = state)) +
  
  # 2. Dynamic foreground layer (participates in animation)
  # Scatter plot and linear regression line
  geom_point(aes(shape = state), alpha = 0.6, size = 4) +
  geom_smooth(method = "lm", se = TRUE, color = "gray50", linewidth = 2, alpha = 0.3,
              show.legend = FALSE) +
  
  # 3. Add cumulative static background shadow map
  shadow_mark(alpha = 0.2, size = 4) +
  
  # 4. optimize chart
  scale_color_manual(values = c("skyblue2", "orange", "pink2", "purple", "yellow3"), name = "state") +
  scale_x_continuous(breaks = seq(0, 100, by = 10), limits = c(0, 100)) +    # x-axis scale and range
  labs(title = "Education Level: {closest_state}",
       subtitle = "Higher education → Lower poverty rate",
       x = "Population with this education level (%)",
       y = "Poverty Rate (%)") +
  theme_minimal() +
  theme(plot.title = element_text(size = 18, face = "bold"),
        plot.subtitle = element_text(size = 14),
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 14),
        panel.border = element_rect(),     # add outer border
        axis.ticks = element_line(),     # add x and y axis ticks
        legend.position = "right",
        legend.title = element_text(size = 15), 
        legend.text = element_text(size = 15)) +
  
  # 5. Display by groups according to categorical variables,
  # showing the poverty rate distribution of 5 states in sequence
  transition_states(edu_type, transition_length = 2, state_length = 1.5) +
  ease_aes('cubic-in-out')    # natural smooth transition

# 6. show animate
animated_plot
```