---
title: "Assignment II: Group B"
author: "WING, Yui Yan"
format: 
  html:
    toc: true
    code-fold: false
    embed-resources: true # <= check here: https://quarto.org/docs/reference/formats/html.html
execute:
  echo: true    # <= knitr::opts_chunk$set(echo = TRUE) in Rmarkdown
  warning: false
  message: false
---

# Name

[**Name**]{style="color:red"}**: WING,Yui Yan**

# Generative AI declaration

Students may use generative AI tools during assignment, for example, to check and improve code, provided that the content is original and reflects the student's own contributions. If generative AI is used, students must declare its use in the box below. An example statement is: “*During the assignment, I used \[XXXX\] to \[XXXX\]. I subsequently reviewed and edited the code and take(s) full responsibility for the final version of the code.”*

::: callout
xxxxxxxxxxxxxx
:::

# Dataset

Each group member may choose either of the following dataset:

**Built-in dataset**

-   `palmerpenguins::penguins`
-   `ggplot2::msleep`
-   `ggplot2::midwest`
-   `datasets::quakes`

**Public dataset**

BIOTIME (CC0) <https://biotime.st-andrews.ac.uk/usageGuidelines.php>

-   `BIOTIME_raw_data_192.csv` (uploaded on Canvas) (Whale data from <https://biotime.st-andrews.ac.uk/selectStudy.php?study=192>)
-   `BIOTIME_raw_data_232.csv` (uploaded on Canvas) (Fish data from <https://biotime.st-andrews.ac.uk/selectStudy.php?study=232>)

# Figures by each group member

### Dataset

::: callout-important
[Which dataset did you choose?]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

xxxxxxxxxxxxxx
:::

### Introduction

::: callout-important
[What is the dataset about? What variables are included in the data? How was the data taken/measured?]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

xxxxxxxxxxxxxx
:::

### Interactive plot(s) using `plotly`

::: callout-important
[Please explain which plot type you choose and why. Also, please discuss the strengths and weaknesses of your plot(s).]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

xxxxxxxxxxxxxx
:::

Note: At least one plot is required, but multiple plots are allowed

::: callout-important
[What is the purpose of this visualization?]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

xxxxxxxxxxxxxx
:::

Note: If you want to specify setting for a code chunk, you can use `#|` for each option (for example, `#| echo: true`). Rmarkdown code chunk still works (for example, `{r echo = true}`), but Quarto style is recommended.

```{r}
#| echo: true
library(tidyverse)   # dplyr, ggplot2, tidyr, readr
library(lubridate)   # safe date handling (though we only use year)
library(plotly)
library(viridis)

whale_data <- read_csv("C:/Users/HP/Downloads/BIOTIME_raw_data_192.csv", show_col_types = FALSE)

# Quick look at structure and first rows
glimpse(whale_data)
head(whale_data)
```
```{r}
library(dplyr) 
whale_clean <- whale_data %>%
  filter(!is.na(ABUNDANCE), ABUNDANCE > 0, !is.na(YEAR)) %>%
  mutate(ABUNDANCE = as.numeric(ABUNDANCE))

# Create a short label for plotting (optional but nicer)
whale_clean <- whale_clean %>%
  mutate(Species_label = case_when(
    str_detect(valid_name, "physalus")    ~ "Fin whale",
    str_detect(valid_name, "musculus")    ~ "Blue whale",
    str_detect(valid_name, "borealis")    ~ "Sei whale",
    str_detect(valid_name, "bonaerensis") ~ "Antarctic minke whale",
    str_detect(valid_name, "novaeangliae")~ "Humpback whale",
    TRUE ~ valid_name   # fallback
  ))


annual_catch <- whale_clean %>%
  group_by(YEAR, Species_label) %>%
  summarise(total_catch = sum(ABUNDANCE, na.rm = TRUE), .groups = "drop")

whale_clean <- whale_data %>%
  filter(!is.na(ABUNDANCE), ABUNDANCE > 0, !is.na(YEAR)) %>%
  # Ensure abundance is numeric (it already is, but safe)
  mutate(ABUNDANCE = as.numeric(ABUNDANCE))

# Check distinct species names (should be 6)
unique(whale_clean$valid_name)


species_list <- unique(annual_catch$Species_label)
n_col <- length(species_list)
base_colors <- viridis(n_col, option = "D")   # n_col 是物种数量

p <- plot_ly(annual_catch, 
             x = ~as.numeric(YEAR), 
             y = ~total_catch, 
             color = ~Species_label,
             colors = base_colors,
             type = 'scatter', 
             mode = 'lines+markers',
             line = list(width = 2),
             marker = list(size = 4),
             text = ~paste("Species:", Species_label, 
                           "<br>Year:", YEAR, 
                           "<br>Catch:", total_catch),
             hoverinfo = "text") %>%
  layout(title = "Annual whale catches (hover to highlight in red)",
         xaxis = list(title = "Year", type = "linear", dtick = 5),
         yaxis = list(title = "Catch"),
         legend = list(title = list(text = "Species")),
         margin = list(b = 60)) %>%
  highlight(on = "plotly_hover",
            dynamic = TRUE,
            selectize = TRUE,
            opacityDim = 0.2)   # 非高亮线条透明度 0.2（数值越小越淡）

p
```





### Animation using `gganimate`

::: callout-important
[Please explain which animation type you choose and why. Also, please discuss the strengths and weaknesses of your animation.]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

xxxxxxxxxxxxxx
:::

Note: At least one animation is required, but multiple animations are allowed

::: callout-important
[What is the purpose of this visualization?]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

xxxxxxxxxxxxxx
:::

```{r}
library(tidyverse)
library(gganimate)
library(gifski)
library(transformr)


whale_clean <- whale_data %>%
  filter(!is.na(ABUNDANCE), ABUNDANCE > 0, !is.na(YEAR)) %>%
  mutate(ABUNDANCE = as.numeric(ABUNDANCE))


whale_clean <- whale_clean %>%
  mutate(Species_label = case_when(
    str_detect(valid_name, "physalus")    ~ "Fin whale",
    str_detect(valid_name, "musculus")    ~ "Blue whale",
    str_detect(valid_name, "borealis")    ~ "Sei whale",
    str_detect(valid_name, "bonaerensis") ~ "Antarctic minke whale",
    str_detect(valid_name, "novaeangliae")~ "Humpback whale",
    TRUE ~ valid_name
  ))


spatial_data <- whale_clean %>%
  filter(!is.na(LATITUDE), !is.na(LONGITUDE)) %>%
  mutate(LONGITUDE_jitter = LONGITUDE + runif(n(), -0.2, 0.2))

# --- 动态计算 y 轴范围 ---
lat_min <- min(spatial_data$LATITUDE, na.rm = TRUE)
lat_max <- max(spatial_data$LATITUDE, na.rm = TRUE)

ylim_custom <- c(lat_min - 1, lat_max + 1)


p_spatial <- ggplot(spatial_data, aes(x = LONGITUDE_jitter, y = LATITUDE, 
                                      size = ABUNDANCE, color = Species_label)) +
  geom_point(alpha = 0.7) +
  scale_size_continuous(range = c(1, 10), name = "Catch (individuals)") +
  scale_color_viridis_d(option = "D", name = "Species") +
  labs(title = "Spatial distribution of whale catches over time",
       subtitle = "Year: {frame_time}",
       x = "Longitude (jittered)", y = "Latitude",
       caption = "Point size = number of individuals caught in 1° cell") +
  coord_quickmap(xlim = c(min(spatial_data$LONGITUDE) - 5, 
                          max(spatial_data$LONGITUDE) + 5),
                 ylim = ylim_custom) +   
  theme_minimal() +
  theme(legend.position = "bottom")


anim_spatial <- p_spatial +
  transition_time(YEAR) +
  ease_aes('linear') +
  shadow_mark(alpha = 0.2, size = 0.8)

anim_spatial
```
