---
title: "Assignment II: Group B"
author: "CHEN, Hongxing"
format: 
  html:
    toc: true
    code-fold: false
    embed-resources: true # <= check here: https://quarto.org/docs/reference/formats/html.html
execute:
  echo: true    # <= knitr::opts_chunk$set(echo = TRUE) in Rmarkdown
  warning: false
  message: false
---

# Name

[**Name**]{style="color:red"}**: CHEN, Hongxing**

# Generative AI declaration

Students may use generative AI tools during assignment, for example, to check and improve code, provided that the content is original and reflects the student's own contributions. If generative AI is used, students must declare its use in the box below. An example statement is: “*During the assignment, I used \[XXXX\] to \[XXXX\]. I subsequently reviewed and edited the code and take(s) full responsibility for the final version of the code.”*

::: callout
xxxxxxxxxxxxxx
:::

# Dataset

Each group member may choose either of the following dataset:

**Built-in dataset**

-   `palmerpenguins::penguins`
-   `ggplot2::msleep`
-   `ggplot2::midwest`
-   `datasets::quakes`

**Public dataset**

BIOTIME (CC0) <https://biotime.st-andrews.ac.uk/usageGuidelines.php>

-   `BIOTIME_raw_data_192.csv` (uploaded on Canvas) (Whale data from <https://biotime.st-andrews.ac.uk/selectStudy.php?study=192>)
-   `BIOTIME_raw_data_232.csv` (uploaded on Canvas) (Fish data from <https://biotime.st-andrews.ac.uk/selectStudy.php?study=232>)

# Figures by each group member

### Dataset

::: callout-important
[Which dataset did you choose?]{style="color:darkred; font-weight:bold;"}
:::

:::: callout
### Your answer is here

I chose the built-in dataset: **palmerpenguins::penguins**.

::: callout
The raw data were obtained from the Environmental Data Initiative (EDI) Data Portal, then lightly curated to standardize variable names and structure, and published under a CC0 public‑domain license (Horst et al., 2022).
:::
::::

```{r}
#| echo: true
# Download the Dataset & Packages required 
install.packages(c("palmerpenguins", "tidyverse", "plotly", "gganimate", "gifski"))
```

```{r}
#| echo: true
# Call the Dataset & Packages used later
library(palmerpenguins)  
library(tidyverse)       
library(plotly)          
library(gganimate)       
library(gifski)          
```

```{r}
#| echo: true
# Load the original dataset & Handle the NaN values
data(penguins)
penguins_clean <- penguins %>% drop_na()

# Show the first 5 rows of the preprocessed dataset
head(penguins_clean)
```

### Introduction

::: callout-important
[What is the dataset about? What variables are included in the data? How was the data taken/measured?]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

This dataset is the **penguins** dataset from the **palmerpenguins** R package, which contains body size measurements for three species of *Pygoscelis* penguins (Adélie, Chinstrap, or Gentoo) that breed on islands throughout the Palmer Archipelago, Antarctic. (Horst et al., 2022). It was created as a modern, real‑world ecological dataset for data visualization and statistical learning, with direct relevance to polar climate change and marine predator research.

The dataset includes **8 variables** for 344 individual penguin observations:

-   **species**: Penguin species (Adélie, Chinstrap, or Gentoo)

-   **island**: Sampling island in the Palmer Archipelago (Biscoe Point, Dream, or Torgersen)

-   **bill_length_mm**: Length of the dorsal ridge of the penguin’s bill (millimeters)

-   **bill_depth_mm**: Depth of the penguin’s bill (millimeters)

-   **flipper_length_mm**: Length of the penguin’s flipper in integear (millimeters)

-   **body_mass_g**: Body mass of the penguin (grams)

-   **sex**: Sex of the penguin (male or female)

-   **year**: Year of observation (2007, 2008, or 2009)

Data were collected from 2007 to 2009 by Dr. Kristen B. Gorman in partnership with the Palmer Station Long Term Ecological Research (LTER) program, part of the U.S. LTER Network (Horst et al., 2022). Wild adult penguins were humanely captured, and standard morphological measurements were taken following formal ecological field protocols. Sex was determined using molecular genetic analysis, and sampling location and year were recorded for each individual.
:::

::: callout
**Citation**: Horst, A., Hill, A. P., & Gorman, K. B. (2022). Palmer Archipelago Penguins Data in the palmerpenguins R Package: An Alternative to Anderson’s Irises. *The R Journal*, 14(1), 244–254. https://doi.org/10.32614/RJ-2022-020
:::

### Interactive plot(s) using `plotly`

::: callout-important
[Please explain which plot type you choose and why. Also, please discuss the strengths and weaknesses of your plot(s).]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

**For** **Plot I**, I chose an **interactive scatter plot with linear regression lines** created by `plotly`. This explanatory plot visualizes the relationship between bill length and flipper length of different types of penguins. In details, points are coded by penguin species (color) and sex (marker shape). I also added species-specific linear regression lines to better understand whether there are within-group linear trends existing. Compared to static plots, this plot supports real-time data interation rather than passive viewing: users can zoom, pan and hover over points to inspect full individual measurement. All in all, this type of plot is ideal for interactively exploring correlations between morphological traits while comparing groups.

**Strengths and weaknesses**

-   **Strengths**: The plot supports interactive functions including zooming, panning and custom hover tooltips that reveal detailed morphological information. Most importantly, the dual-legend design allows independent clicking of legend items to show or hide specific groups, letting users filter out clutter and focus on targeted comparisons. This level of focused analysis is impossible in static versions. In addition, the linear trends clearly show wthether the within-specific correlations existed.

-   **Weaknesses**: The plot is restricted to a fixed pair of morphological features. The users cannot switch to other body measurements. What's more, the overlapping issues appears in dense data regions. It's hard to distinguish which types of point it is. The `island` and `year` variables are hard to be used in such plots. It cannot show the correlation in accurate numbers.\

**For** **Plot II**, I chose a box plot overlaid with jitter raw data points to display the distribution of body mass across penguin species and sex. Box plot is ideal for this explanatory visualization, as it can clearly summarize statistical distributions (median, IQR, outliers) for group comparisons. I also added jitter points to display every individual penguin’s body mass, which complements the aggregated box plot statistics with raw data context. While the interactive legend allows viewers to toggle sex groups for communicating focused biological patterns.

**Strengths and Weaknesses**

-   **Strengths**: The plot effectively highlights two key biological trends: (1) interspecific size divergence (Gentoo penguins are substantially heavier than Adelie and Chinstrap), and (2) sexual dimorphism (male penguins are consistently larger than females within each species). In addition, the Plotly legend allows viewers to toggle visibility of sexes, making it easy to isolate and compare groups for a deeper understanding of body mass variation. Last but not least, by combining box plot statistics with individual data points, the plot provides both a high-level summary and granular detail about the population.

-   **Weaknesses**: The plot aggregates data across all three islands, so it does not show potential variation in body mass across different breeding sites. Despite jitter, individual data points for larger groups (e.g., male Gentoo penguins) still overlap slightly, which may obscure fine-grained variation.
:::

Note: At least one plot is required, but multiple plots are allowed

::: callout-important
[What is the purpose of this visualization?]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

**For Plot I**, the purpose is to clearly **obtain the linear relationship between penguin bill length and flipper length across different species and sexes.** Obviously, the **Gentoo penguin has a more significant positive relationship between penguin bill length and flipper length (in both male or female) than other penguins.** The other two penguins' correlation still suffer from many outliers. Furthermore, this visualization also communicate both intraspecific and interspecific morphological differences, and we can clearly figure out 3 types of clusters.

**For Plot II**, the purpose is to **explain the significant differences in body mass distribution between penguin species and sexes**. Obviously, the universal trend of sexual dimorphism exists since **male penguins are consistently larger than females within each species**. Meanwhile, the **Gentoo penguins have the largest body mass**. This plot provides clear evidence of morphological divergence between species.
:::

Note: If you want to specify setting for a code chunk, you can use `#|` for each option (for example, `#| echo: true`). Rmarkdown code chunk still works (for example, `{r echo = true}`), but Quarto style is recommended.

### Plot I : Relationship between Bill Length and Flipper Length for Penguin

```{r}
#| echo: true
#| warning: false
#| message: false

library(palmerpenguins)
library(tidyverse)
library(plotly)

# 数据清洗
penguins_clean <- penguins %>% drop_na()

# 定义颜色和形状映射
species_colors <- c("Adelie" = "#1f77b4", "Chinstrap" = "#ff7f0e", "Gentoo" = "#2ca02c")
sex_symbols   <- c("female" = "circle", "male" = "triangle-up")

# 按 species + sex 分组
groups <- split(penguins_clean, 
                list(penguins_clean$species, penguins_clean$sex), 
                drop = TRUE)

# 初始化 plotly 对象
p <- plot_ly()

# ---------- 添加散点 traces（属于 legend 图例） ----------
for (nm in names(groups)) {
  g  <- groups[[nm]]
  sp <- as.character(unique(g$species))
  sx <- as.character(unique(g$sex))

  p <- add_trace(
    p,
    data   = g,
    x      = ~bill_length_mm,
    y      = ~flipper_length_mm,
    type   = "scatter",
    mode   = "markers",
    name   = paste(sp, sx),                 # 图例显示 "Adelie female" 等
    legend = "legend",                      # 指定属于第一个图例
    marker = list(
      size    = 7,
      opacity = 0.8,
      color   = species_colors[[sp]],       # 直接指定颜色
      symbol  = sex_symbols[[sx]],          # 直接指定符号
      line    = list(width = 0.8, color = "white")
    ),
    text = ~paste(
      "Species:", species,
      "\nSex:", sex,
      "\nIsland:", island,
      "\nBill Length:", bill_length_mm, "mm",
      "\nFlipper Length:", flipper_length_mm, "mm",
      "\nBody Mass:", body_mass_g, "g"
    ),
    hoverinfo = "text"
  )
}

# ---------- 添加回归线 traces（属于 legend2 图例） ----------
# 为每个物种分别拟合线性回归，保证线条连续且颜色正确
for (sp in c("Adelie", "Chinstrap", "Gentoo")) {
  sp_data <- penguins_clean %>% filter(species == sp)
  model   <- lm(flipper_length_mm ~ bill_length_mm, data = sp_data)
  
  # 生成用于绘制回归线的 x 序列和预测值
  x_range <- range(sp_data$bill_length_mm, na.rm = TRUE)
  x_seq   <- seq(x_range[1], x_range[2], length.out = 50)
  y_pred  <- predict(model, newdata = data.frame(bill_length_mm = x_seq))

  p <- add_trace(
    p,
    x       = x_seq,
    y       = y_pred,
    type    = "scatter",
    mode    = "lines",
    name    = sp,                           # 图例只显示物种名
    legend  = "legend2",                    # 指定属于第二个图例
    line    = list(color = species_colors[[sp]], width = 2),
    hoverinfo = "skip"
  )
}


p %>%
  layout(
    title = list(
      text = "Relationship between Bill Length and Flipper Length for Penguin",
      x    = 0.05,
      y    = 0.88
    ),
    xaxis = list(title = "Bill Length (mm)"),
    yaxis = list(title = "Flipper Length (mm)"),
    # 第一个图例：散点（物种 + 性别）
    legend = list(
      title = list(text = "Penguin Species & Sex"),
      x     = 1.02,
      y     = 1,
      xanchor = "left"
    ),
    # 第二个图例：拟合线
    legend2 = list(
      title = list(text = "Fitted Line"),
      x     = 1.02,
      y     = 0.1,
      xanchor = "left"
    ),
    plot_bgcolor  = "white",
    paper_bgcolor = "white",
    margin = list(r = 200, l = 60, t = 80, b = 60)
  )
```

### Plot II : Body Mass Distribution by Species and Sex

```{r}
#| echo: true
library(palmerpenguins)
library(tidyverse)
library(plotly)

penguins_clean <- penguins %>% drop_na()

plot_ly(penguins_clean, 
        x = ~species, 
        y = ~body_mass_g, 
        color = ~sex,
        colors = c("female" = "pink", "male" = "darkblue"),
        type = "box",
        boxpoints = "all",
        jitter = 0.3,
        pointpos = -1.8) %>%
  layout(
    title = "Body Mass Distribution by Species and Sex",
    xaxis = list(title = "Penguin Species"),
    yaxis = list(title = "Body Mass (g)"),
    legend = list(
      title = list(text = "Sex"),
      x = 1,           # 将图例放置在绘图区域右侧外部
      y = 1,
      xanchor = "left",
      yanchor = "top"
    ),
    margin = list(l = 60, r = 30, t = 60, b = 80) 
  )
```

::: callout
### Extra Exploratory Plot due to personal interest (Can Skip)

For Plot III, this is an **enhanced explanatory interactive scatter plot** with a default view and optional axis-switching dropdown menus. By default, it displays the clear explanatory relationship between body mass and flipper length — structurally and conceptually similar to Plot 1. But, the dropdown menus are additional supportive features to let users access other meaning trait relationships within the same figure, rather than requiring multiple separate explanatory plots.

Strengths and weaknesses

-   Strengths: In its default state, it functions as a clean and focused explanatory visualization, like Plot I, to efficiently communicating the positive relationship between the body mass and flipper length among 3 types of penguin. The added dropdown menus bring practical convenience by allowing users exploring other interesting relationships without switching figures. Legend item clicking still allows users to isolate individual species-sex groups, removing unrelated data to reveal subtle patterns. Standard Plotly interactions (zoom, pan, or hover) further support detailed inspection. Thus, this plot consolidates high-diemnsional data exploration into a single interactive visualization.

-   Weaknesses: No regression lines are included to highlight linear trends. Similarly, the overlapping issues still appears in dense data regions. Although the plot has included most of features from the data, it still misses the `island` and `year` variables. Viewers must actively interact to access other potential relationships, which are not immediately visible,

For Plot III, the plot primarily serves to **clearly explain the positive relationship between body mass and flipper length** through its default view. (i.e., by default selecting body mass and flipper length, the Adelie and Chinstrap penguins have a closer relationship than that of the Gentoo penguins, meanwhile these 3 types of penguin has the positive relationships between these two features). On top of this core function, the dropdown menus are included to let interested viewers conveniently examine other potential morphological relationships within the same figure, eliminating the need to create and consult multiple separate plots for different trait comparisons.
:::

### Plot III - Exploratory Plot for Penguin Morphological Traits (Can Skip)

```{r}

#| echo: true
#| warning: false
#| message: false

library(palmerpenguins)
library(plotly)

penguins_clean <- na.omit(penguins)

# 按 species + sex 组合手动分组
groups <- split(penguins_clean, 
                list(penguins_clean$species, penguins_clean$sex), 
                drop = TRUE)

species_colors <- c(
  "Adelie"    = "#1f77b4",
  "Chinstrap" = "#ff7f0e",
  "Gentoo"    = "#2ca02c"
)
sex_symbols <- c("female" = "circle", "male" = "triangle-up")

# 初始轴变量（默认 X = Body Mass, Y = Flipper Length）
default_x <- "body_mass_g"
default_y <- "flipper_length_mm"

# ---- 构建初始图：每个 species-sex 组合作为一个独立 trace ----
p <- plot_ly()

for (nm in names(groups)) {
  g  <- groups[[nm]]
  sp <- as.character(unique(g$species))
  sx <- as.character(unique(g$sex))
  
  p <- add_trace(
    p,
    x           = g[[default_x]],
    y           = g[[default_y]],
    type        = "scatter",
    mode        = "markers",
    name        = paste(sp, sx),          # 图例显示 "Adelie female" 等
    # 不设置 legendgroup，或设置为唯一值，确保每个 trace 独立
    legendgroup = paste(sp, sx),
    showlegend  = TRUE,
    marker      = list(
      size    = 8,
      opacity = 0.85,
      color   = species_colors[[sp]],
      symbol  = sex_symbols[[sx]],
      line    = list(width = 0.8, color = "white")
    ),
    text        = paste(
      "<b>Species:</b>", g$species, "<br>",
      "<b>Sex:</b>", g$sex, "<br>",
      "<b>Island:</b>", g$island, "<br>",
      "<b>Body Mass:</b>", g$body_mass_g, "g<br>",
      "<b>Bill Length:</b>", g$bill_length_mm, "mm<br>",
      "<b>Bill Depth:</b>", g$bill_depth_mm, "mm<br>",
      "<b>Flipper Length:</b>", g$flipper_length_mm, "mm"
    ),
    hoverinfo   = "text"
  )
}

# 生成 X 轴菜单按钮
make_x_button <- function(col_name, x_title, label) {
  x_vals <- lapply(names(groups), function(nm) groups[[nm]][[col_name]])
  list(
    method = "update",
    label  = label,
    args   = list(
      list(x = x_vals),
      list(xaxis = list(title = x_title))
    )
  )
}

# 生成 Y 轴菜单按钮
make_y_button <- function(col_name, y_title, label) {
  y_vals <- lapply(names(groups), function(nm) groups[[nm]][[col_name]])
  list(
    method = "update",
    label  = label,
    args   = list(
      list(y = y_vals),
      list(yaxis = list(title = y_title))
    )
  )
}

# 添加布局与下拉菜单
p %>%
  layout(
    title = list(
      text = "Interactive Comparison of Penguin Morphological Traits",
      x    = 0.05,      
      y    = 0.9       
    ),
    xaxis = list(title = "Body Mass (g)"),
    yaxis = list(title = "Flipper Length (mm)"),
    hovermode = "closest",
    margin = list(r = 180, l = 60, t = 80, b = 60),
    # 图例设置）
    legend = list(
      title = list(text = "Penguin Species & Sex"),
      x = 1.02, y = 1.05,
      xanchor = "left"
    ),
    # 给两个下拉菜单添加标题
    annotations = list(
      # X轴下拉菜单标题
      list(
        x = 1.15, y = 0.2,
        xref = "paper", yref = "paper",
        text = "X-axis",
        showarrow = FALSE,
        font = list(size = 11, color = "black")
      ),
      # Y轴下拉菜单标题
      list(
        x = 1.15, y = 0.0,
        xref = "paper", yref = "paper",
        text = "Y-axis",
        showarrow = FALSE,
        font = list(size = 11, color = "black")
      )
    ),
    updatemenus = list(
      # X 轴下拉菜单
      list(
        x       = 1.15,          
        y       = 0.3,
        xanchor = "left",
        yanchor = "top",
        active  = 3,             
        buttons = list(
          make_x_button("bill_length_mm",   "Bill Length (mm)",   "Bill Length"),
          make_x_button("bill_depth_mm",    "Bill Depth (mm)",    "Bill Depth"),
          make_x_button("flipper_length_mm","Flipper Length (mm)","Flipper Length"),
          make_x_button("body_mass_g",      "Body Mass (g)",      "Body Mass")
        )
      ),
      # Y 轴下拉菜单
      list(
        x       = 1.15,
        y       = 0.1,
        xanchor = "left",
        yanchor = "top",
        active  = 2,             
        buttons = list(
          make_y_button("bill_length_mm",   "Bill Length (mm)",   "Bill Length"),
          make_y_button("bill_depth_mm",    "Bill Depth (mm)",    "Bill Depth"),
          make_y_button("flipper_length_mm","Flipper Length (mm)","Flipper Length"),
          make_y_button("body_mass_g",      "Body Mass (g)",      "Body Mass")
        )
      )
    )
  )
```

### Animation using `gganimate`

::: callout-important
[Please explain which animation type you choose and why. Also, please discuss the strengths and weaknesses of your animation.]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

**For Animation I**, this is an **explanatory animated bar chart** that shows penguin population counts and species composition across Biscoe, Dream and Torgersen Islands from 2007 to 2009. Each bar is labelled with the percentage of each species relative to the total island population in a given year. The animated transition over years and faceted layout help viewers clearly observe both population size and proportional structure in a smooth and intuitive way.

**Strengths and Weaknesses**

-   **Strengths**: The smooth animation clearly reveals that penguin habitat distribution and species proportions remain highly stable across the three years. Percentage labels directly show species composition without extra calculation, improving interpretability. Faceting by island supports straightforward spatial comparison, and the consistent colour scheme maintains visual coherence with other plots.

-   **Weaknesses**: Annual changes in population and proportion are small, so the animation effect is relatively gentle. The chart presents aggregated yearly data and does not reflect individual-level variation. Also, the plot ignores the sex of the penguins.

**For Animation II**, this is an **explanatory animated boxplot**, focusing exclusively on the body mass distribution of Adelie penguins across Biscoe, Dream and Torgersen Islands from 2007 to 2009. A boxplot is chosen as it effectively summarizes key statistical features of Adelie body mass (median, interquartile range, and outliers) across different islands, while the year-by-year animation reveals subtle temporal changes in body mass distribution. The faceted layout by island allows for clear comparison of Adelie body mass across their habitats,

**Strengths and Weaknesses**

-   **Strengths**: Focuses solely on Adelie penguins, providing a targeted analysis of their body mass stability across three islands and three years. The boxplot clearly displays the distribution and central tendency of Adelie body mass, while the animation smoothly illustrates minor annual fluctuations without overwhelming the viewer. Effectively complements Animation I by linking population dynamics to morphological stability, strengthening the overall conclusion of stable Adelie colonies.

-   **Weaknesses**: The animation effect is relatively subtle due to the small annual changes in Adelie body mass, making dramatic temporal shifts less apparent. The plot does not include a sex dimension, so it fails to reveal potential sexual dimorphism in Adelie body mass across islands.

**For Animation III**, this is an **explanatory animated boxplot**, illustrating the body mass distribution of male and female Adelie penguins across Biscoe, Dream and Torgersen Islands from 2007 to 2009. The boxplot is selected because it effectively visualises the central tendency and spread of body mass separately for male and female Adelie penguins, making sexual size dimorphism easy to observe. Animation over years is used to reveal temporal changes in body mass across sexes, while faceting by island supports direct habitat comparison. This plot also extends the previous morphological animation by introducing sex as a key variable.

**Strengths and Weaknesses**

-   **Strengths**: The boxplot clearly separates body mass distributions by sex, allowing straightforward detection of sexual dimorphism without visual distortion. Animated transitions smoothly display year‑to‑year shifts in body mass, preserving continuity with the previous animation. In additioon, faceting by island improves readability and enables side‑by‑side comparison across the three habitats.

-   **Weaknesses**: Some annual changes in body mass are relatively small, resulting in subtle animation that may not immediately draw attention. The plot focuses only on Adelie penguins and provides no interspecific reference for context. Without direct numerical labels on the boxplots, viewers must rely on visual judgment to compare exact values.
:::

Note: At least one animation is required, but multiple animations are allowed

::: callout-important
[What is the purpose of this visualization?]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

**For Animation I**, this animation aims to **reveal the penguin species-habitat distribution and species population** **across the three islands over 2007 to 2009.** It highlights the fixed species-habitat segregation over years: **Gentoo penguins found in Bisco Island only; Chinstrap penguins found in Dream Island only; Adelie penguins found in all 3 islands**. It further illustrates the subtle year-to-year fluctuations in population proportions and counts on each island, including the **dominant share of Gentoo penguins in Biscoe** with temporary narrowing of species gaps, **mild shifts in Chinstrap dominance in Dream**, and the steady **slow growth of the sole Adelie population in Torgersen**, demonstrating both persistent habitat preferences and gentle temporal dynamics in penguin colony sizes.

**For Animation II**, this animated boxplot **reveal the** **body mass distribution of Adelie penguins across the three islands over 2007 to 2009**. It illustrates the specific temporal fluctuations in Adelie body mass: **Torgersen Island had the heaviest Adelie penguins in 2007 and 2008**, all three islands saw an increase in body mass in 2008. While **in 2009, Biscoe Island’s Adelie penguins rose to the heaviest**, and Dream and Torgersen Islands experienced a decline. This **animation complements the population dynamics shown in Animation I by linking the annual changes in Adelie penguin population to their corresponding body mass variations**, revealing how Adelie body mass changed alongside their population fluctuations each year and providing clear insights into the body mass status of Adelie penguins during each period of population change.

**For Animation III**, this animation **reveals body mass patterns between male and female Adelie penguins across the three Antarctic islands from 2007 to 2009**. It confirms strong and consistent sexual dimorphism, with **males being noticeably heavier than females across all islands and years**. It also reproduces the same temporal pattern observed in Animation II: **both male and female Adelie penguins on Torgersen Island were the heaviest during 2007 and 2008**, whereas b**oth sexes on Biscoe Island became the heaviest in 2009**. This animation addresses the limitation of Animation II by including sex‑based differentiation and testing whether the observed body mass shifts were influenced by gender. By showing consistent patterns across both sexes, it **reinforces the reliability of the island‑specific temporal trends and complements the population dynamics shown in Animation II**, completing a more holistic analysis of Adelie penguin.
:::

### Animation I : Penguin Habitat Distribution Across Three Islands (2007-2009)

```{r}
#| echo: true
#| warning: false
#| message: false

library(palmerpenguins)
library(tidyverse)
library(gganimate)

# 数据准备
penguins_clean <- penguins %>% drop_na()

penguin_distribution <- penguins_clean %>%
  group_by(year, island, species) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(year, island) %>%
  mutate(percent = count / sum(count) * 100) %>%
  ungroup() %>%
  mutate(
    year = as.numeric(year),
    group_id = paste(island, species)
  )

# 动画代码
anim1 <- ggplot(
  penguin_distribution,
  aes(x = species, y = count, fill = species, group = group_id)
) +
  geom_col(alpha = 0.8, width = 0.6) +
  
  geom_text(
    aes(label = sprintf("%.1f", percent)),     
    vjust = -1,
    size = 5,
    fontface = "bold",
    color = "black"
  ) +
  
  geom_text(
    aes(label = "%"),
    vjust = -1,
    size = 5,
    fontface = "bold",
    color = "black",
    hjust = 0,
    position = position_nudge(x = 0.22)
  ) +
  
  facet_wrap(~island) +
  scale_fill_manual(values = c(
    "Adelie" = "#1f77b4",
    "Chinstrap" = "#ff7f0e",
    "Gentoo" = "#2ca02c"
  )) +
  labs(
    title = "Penguin Habitat Distribution Across Three Islands (2007-2009)",
    subtitle = "Year: {round(frame_time, 0)}",
    x = "Species",
    y = "Number of Individuals",
    fill = "Species"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 16, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold", margin = margin(r = 15)),
    axis.title.x = element_text(size = 14, face = "bold", margin = margin(t = 10)),
    axis.text = element_text(size = 13, face = "bold"),
    strip.text = element_text(size = 14, face = "bold"),
    legend.title = element_text(size = 14, face = "bold"),
    legend.text = element_text(size = 13, face = "bold")
  ) +
  transition_time(year) +
  ease_aes('sine-in-out')

animate(anim1, 
        nframes = 200, 
        fps = 20, 
        width = 950, 
        height = 600,
        end_pause = 30)
```

### Animation II : Adelie Penguin Body Mass Distribution Across Three Islands (2007-2009)

```{r}
#| echo: true
#| warning: false
#| message: false

library(palmerpenguins)
library(tidyverse)
library(gganimate)

# 数据清洗：仅保留 Adelie 企鹅）
penguins_clean <- penguins %>% drop_na() %>%
  filter(species == "Adelie") %>%  # 
  mutate(year = as.numeric(year))

# 动画绘图：
anim2 <- ggplot(penguins_clean, 
                aes(x = species, y = body_mass_g, fill = species)) +
  geom_boxplot(alpha = 0.8, outlier.size = 2) +
  
  facet_wrap(~island) +
  scale_fill_manual(values = c("Adelie" = "#1f77b4")) +
  labs(
    title = "Adelie Penguin Body Mass Distribution Across Three Islands (2007-2009)",
    subtitle = "Year: {round(frame_time, 0)}",
    x = "Species",
    y = "Body Mass (g)",
    fill = "Species",
    caption = "Adelie body mass remains stable across three islands\nwith minor annual fluctuations"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 16, face = "bold"),
    plot.caption = element_text(size = 12, face = "bold", hjust = 1, lineheight = 1.2, margin = margin(t = 15, b = 5)),
    axis.title.y = element_text(size = 14, face = "bold", margin = margin(r = 15)),
    axis.title.x = element_text(size = 14, face = "bold", margin = margin(t = 10)),
    axis.text = element_text(size = 13, face = "bold"),
    strip.text = element_text(size = 14, face = "bold"),
    legend.title = element_text(size = 14, face = "bold"),
    legend.text = element_text(size = 13, face = "bold")
  ) +
  transition_time(year) +
  ease_aes('sine-in-out')

animate(anim2, 
        nframes = 200, 
        fps = 20, 
        width = 950, 
        height = 580,
        end_pause = 30)
```

### Animation III : Adelie Penguin Body Mass by Sex Across Three Islands (2007-2009)

```{r}
#| echo: true
#| warning: false
#| message: false

library(palmerpenguins)
library(tidyverse)
library(gganimate)

# 仅 Adelie 企鹅 + 清洗性别数据
penguins_clean <- penguins %>% drop_na() %>%
  filter(species == "Adelie") %>%
  mutate(year = as.numeric(year))

anim3 <- ggplot(penguins_clean, 
                aes(x = sex, y = body_mass_g, fill = sex)) +
  geom_boxplot(alpha = 0.8, outlier.size = 2) +
  
  facet_wrap(~island) +
  scale_fill_manual(values = c("male" = "#4472C4", "female" = "#E0578F")) +
  labs(
    title = "Adelie Penguin Body Mass by Sex Across Three Islands (2007-2009)",
    subtitle = "Year: {round(frame_time, 0)}",
    x = "Sex",
    y = "Body Mass (g)",
    fill = "Sex"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 16, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold", margin = margin(r = 15)),
    axis.title.x = element_text(size = 14, face = "bold", margin = margin(t = 10)),
    axis.text = element_text(size = 13, face = "bold"),
    strip.text = element_text(size = 14, face = "bold"),
    legend.title = element_text(size = 14, face = "bold"),
    legend.text = element_text(size = 13, face = "bold")
  ) +
  transition_time(year) +
  ease_aes('sine-in-out')

animate(anim3, 
        nframes = 200, 
        fps = 20, 
        width = 950, 
        height = 580,
        end_pause = 30)
```
